---
title: "Working with sf objects"
author: "Sofie Lloyd Thomsen"
date: "Created 04-February-2022 updated `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

#  Goals: 

1. In the first exercise, you will learn to construct maps with tmaps() by modifying the provided code. The libraries, data, and initial code is provided below.
2. In the second exercise, you learn to: 
 - Understand the provided datasets
 - Learn how to reproject spatial data
 - Limit your data into an area of interest
 - Create a new map
3. In the third exercise, you reproject layers to a shared SRS.
4. In the fourth (and optional) exercise, you create a map of Denmark using online data.

#### Required R libraries

We will use the sf, raster, and tmap packages.
Additionally, we will use the spData and spDataLarge packages that provide new datasets.

```{r}
library(sf)
library(raster)
library(tmap)
library(spData)

# To install the spDataLarge uncomment the next line
install.packages("spDataLarge", repos = "https://geocompr.r-universe.dev")
library(spDataLarge)
```


#  Exercise 1 - Make a map

In this exercise you will learn to make a map with tmaps library, by adding spatial data layers and modifying the arguments that specify their rendering

#### Data sets  

We will use two data sets: `nz_elev` and `nz`. They are contained by the libraries
The first one is an elevation raster object for the New Zealand area, and the second one is an sf object with polygons representing the 16 regions of New Zealand.

#### Existing code

Here is the code to create a new map of New Zealand (following Robin Lovelace's example).Your role is to improve this map based on the suggestions below.

```{r initial-map}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "BuGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "A. Sobotkova, 2022") +
  tm_layout(main.title = "My map",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```


## Tasks 

1. Change the map title from "My map" to "New Zealand".
2. Update the map credits with your own name and today's date.
3. Change the color palette to "-RdYlGn". 
  (You can also try other palettes from http://colorbrewer2.org/)
4. Put the north arrow in the top right corner of the map.
5. Improve the legend title by adding the used units (m asl).
6. Increase the number of breaks in the scale bar.
7. Change the borders' color of the New Zealand's regions to black. 
  Decrease the line width.
8. Change the background color to any color of your choice.


## Your solution

```{r improved-map}
# /Start Code/
tm_shape(nz_elev)  +
  
# Improve the legend title by adding the used units (m asl)
  tm_raster(title = "elev (m asl)", 
            style = "cont",
            
          # Changing the color palette to "-RdYlGn"
            palette = "-RdYlGn") +
  tm_shape(nz) +
  
# Change the borders' color of the New Zealand's regions to black
  tm_borders(col = "black", 
             
            # Decrease the line width
             lwd = 1) +
  
# Increase the number of breaks in the scale bar
  tm_scale_bar(breaks = c(0, 100, 200, 300, 400),
               
              # decrease text size
               text.size = 0.5) +
  
# Put the north arrow in the top right corner of the map
  tm_compass(position = c("RIGHT", "TOP"),
             type = "rose", 
             size = 2) +
  
# Update the map credits with my own name and today's date
  tm_credits(text = "Sofie L. Thomsen, 09/02/2022",
             size = 0.5) + 
  
# Changing the main title to 'New Zealand'
  tm_layout(main.title = "New Zealand", 
            
          # Change the background color to any color of your choice.
            bg.color = "light blue",
            inner.margins = c(0, 0, 0, 0))

# /End Code/

```


# Exercise 2 - Inspect and display spatial data

We will use two data sets: `srtm` and `zion` , explore them and make a map.
The first one is an elevation raster object for the Zion National Park area in Israel, and the second one is an sf object with polygons representing borders of the Zion National Park.

Read two new datasets, `srtm` and `zion`, using the code below and then follow the instructions and answer the questions related.

####  Preparation: Load data from provided package
You need to have installed the spDataLarge package for the following code to work. 

```{r load-data}
srtm  <-  raster(system.file("raster/srtm.tif", package = "spDataLarge"))
zion  <-  read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
```

## Instructions
Type answers to the questions as code comments next to or under the code used

1. Display the `zion` object and view its structure.

2. Display the `srtm` object and view its structure.


## My solution and answers

**Inspecting the 'zion' object**

```{r inspect-zion}
# /Start Code /

# View the structure of the data
str(zion)
summary(zion)

# View coordinate reference system
st_crs(zion)

# View attributes
attributes(zion)

# View geometry
st_geometry(zion)

# /End Code /

```

*Questions and answers*

  - What can you say about the content of this file?  
  The 'zion' dataset consists of a dataframe with 1 observation and 12 variables.  
  <p>&nbsp;</p>
  - What type of data does it store?  
  The datatypes of the data is primarily characters, but also numerics.  
  <p>&nbsp;</p>
  - What is the coordinate system used?  
  The coordinate system that is being used is Urchin Tracking Module (UTM), zone 12
<p>&nbsp;</p> 
  - How many attributes does it contain?  
  The dataset consists of five attributes.
  <p>&nbsp;</p>
  - What is its geometry?
  The 'zion' datasets geometry is polygon
  

**Inspecting the 'strm' object**

```{r inspect-strm}
# /Start Code /

# View the structure of the data
str(srtm)

# View the Coordinate Reference System
st_crs(srtm)

# View attributes
attributes(srtm)

# View dimensions 
dim(srtm)

# View data resolution 
res(srtm)

# /End Code /
```


*Questions and answers*

  - What can you say about the content of this file?  
  The strm dataset consists of raster layers with 12 slots.  
<p>&nbsp;</p>
  - What type of data does it store?  
  The datatypes is logicals, numerics, integers and characters.  
  <p>&nbsp;</p>
  - What is the coordinate system used?  
  The coordinate system that is being used is the "World Geodetic System 1984" (WGS84). 
  <p>&nbsp;</p>
  - How many attributes does it contain?
  The datasets has 13 attributes.
<p>&nbsp;</p>
  - How many dimensions does it have? 
  The dataset has 3 dimensions.  
  <p>&nbsp;</p>
  - What is the data resolution?  
  The data resolution is 0.0008333333 : 0.0008333333



# Exercise 3 - Reconciling Coordinate systems

Now that you know that coordinate systems differ, make them compatible!


## Instructions

* Reproject the `srtm` dataset into the coordinate reference system used in the `zion` object. 

  - Create a new object `srtm2`
  - Visualize the results using the `plot()` function.

* Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object.
  - Create a new object `zion2`
  - Visualize the results (`zion2` together with `srtm2`) using the `plot()` function.

## Your solution

```{r reproject srtm into zion}
# /Start Code/ #

# Reproject 
srtm2 <- projectRaster(srtm,
                       crs = crs(zion))

# Visualise 'srtm2'
plot(srtm2)


# /End Code/ #
```


```{r reproject zion into srtm}
# Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object
zion2 <- st_transform(zion,
                       crs = crs(srtm))

# Visualise 'zion2'
plot(zion2)

```

```{r}
# Visualize the results (`zion2` together with `srtm2`) using the `plot()` function.

# Save the CRS of the srtm2 layer
srtm2_crs <- crs(srtm2, asText = TRUE)

# Transform the zion2 CRS to match srtm2
zion2_crs <- st_transform(zion2, crs = srtm2_crs)

# Run both lines together
plot(srtm2)
plot(zion2_crs, add = TRUE,
     col = "transparent")
```


# Exercise 4 - Make a map of Denmark (Optional)

Make a handsome map of Denmark from data available online with a scale, a north arrow and source credits!

## Hints

*Load elevation and regional boundaries for Denmark using the `getData()` function from the `raster` package. 

  - Remember that the `getData()` function takes a number of different attributes, plus the country code in a ISO 3-letter format. You can google what it is for Denmark, or use the `getData('ISO3')` function to view it. For an example of how to use the `getData()` function [see this blog](https://www.gis-blog.com/r-raster-data-acquisition/)
* Make a map with `tmaps` library following the Exercise no.1

## Your solution
```{r}
# /Start Code/ 

# Load data
Denmark <- getData('GADM', country='DK', level=1)

# Make map
tm_shape(Denmark)  +
  tm_polygons(col = "green") +
  tm_shape(Denmark) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_scale_bar(breaks = c(0, 50, 100, 150, 200),
              text.size = 0.5,
              position = c("RIGHT", "TOP")) +
  tm_compass(position = c("LEFT", "TOP"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "Sofie L. Thomsen, 09/02/2022") + 
  tm_layout(main.title = "Denmark",
            bg.color = "light blue",
            inner.margins = c(0, 0, 0, 0))

# /End Code/ #
```

